---
title: "p8105 hw6 yt2785"
author: Yijing Tao (UNI yt2785)
date: Nov 2021
output: github_document
---

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

set.seed(1)
```

## Problem 1

Read and clean the data.
```{r warning=FALSE}
#read csv
birthweight_df = 
  read_csv("./birthweight.csv") %>% 
  mutate(
    babysex = recode(babysex, '1' = "male", '2' = "female"),
    babysex = as.factor(babysex),
    frace = recode(frace, '1' = "white", '2' = "black", '3' = "asian", '4' = "puerto_rican", '8' = "other", '9' = "unknown"),
    frace = as.factor(frace),
    malform = recode(malform, '0' = "absent", '1' = "present"),
    malform = as.factor(malform),
    mrace = recode(mrace, '1' = "white", '2' = "black", '3' = "asian", '4' = "puerto_rican", '8' = "other"),
    mrace = as.factor(mrace)
  )

#check whether there are missing datas
na_bwt = any(is.na(birthweight_df))
```

Transform `babysex`, `frace`, `malform` and `mrace` variables into factor variables. And the answer of having missing data or not is `r na_bwt`, which means there is no missing data in this data set.

*Make the plots showing all of the relationships between numeric variables and birthweight.*

```{r warning=FALSE}
birthweight_df %>% 
  ggplot(aes(x = bhead, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = blength, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = delwt, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = fincome, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = gaweeks, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = menarche, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = mheight, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = momage, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = parity, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = pnumlbw, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = pnumsga, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = ppbmi, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = ppwt, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = smoken, y = bwt)) + 
  geom_point()

birthweight_df %>% 
  ggplot(aes(x = wtgain, y = bwt)) + 
  geom_point()
```

From the plots above we can infer that the birthweight is most related to `bhead` and `blength`.

*Hypothesis testing*

```{r warning=FALSE}
fit_null1 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt1 = lm(bwt ~ bhead + blength + delwt, data = birthweight_df)

anova(fit_null1, fit_alt1) %>% 
  broom::tidy()

fit_null2 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt2 = lm(bwt ~ bhead + blength + fincome, data = birthweight_df)

anova(fit_null2, fit_alt2) %>% 
  broom::tidy()

fit_null3 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt3 = lm(bwt ~ bhead + blength + gaweeks, data = birthweight_df)

anova(fit_null3, fit_alt3) %>% 
  broom::tidy()

fit_null4 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt4 = lm(bwt ~ bhead + blength + menarche, data = birthweight_df)

anova(fit_null4, fit_alt4) %>% 
  broom::tidy()

fit_null5 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt5 = lm(bwt ~ bhead + blength + mheight, data = birthweight_df)

anova(fit_null5, fit_alt5) %>% 
  broom::tidy()

fit_null6 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt6 = lm(bwt ~ bhead + blength + momage, data = birthweight_df)

anova(fit_null6, fit_alt6) %>% 
  broom::tidy()

fit_null7 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt7 = lm(bwt ~ bhead + blength + parity, data = birthweight_df)

anova(fit_null7, fit_alt7) %>% 
  broom::tidy()

fit_null8 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt8 = lm(bwt ~ bhead + blength + pnumlbw, data = birthweight_df)

anova(fit_null8, fit_alt8) %>% 
  broom::tidy()

fit_null10 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt10 = lm(bwt ~ bhead + blength + ppbmi, data = birthweight_df)

anova(fit_null10, fit_alt10) %>% 
  broom::tidy()

fit_null11 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt11 = lm(bwt ~ bhead + blength + ppwt, data = birthweight_df)

anova(fit_null11, fit_alt11) %>% 
  broom::tidy()

fit_null12 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt12 = lm(bwt ~ bhead + blength + smoken, data = birthweight_df)

anova(fit_null12, fit_alt12) %>% 
  broom::tidy()

fit_null13 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt13 = lm(bwt ~ bhead + blength + wtgain, data = birthweight_df)

anova(fit_null13, fit_alt13) %>% 
  broom::tidy()

fit_null14 = lm(bwt ~ bhead + blength, data = birthweight_df)
fit_alt14 = lm(bwt ~ bhead + blength + pnumsga, data = birthweight_df)

anova(fit_null14, fit_alt14) %>% 
  broom::tidy()
```

Since we can reject the null hypothesis only when p.value is extremly small,  from the anova outcome above we can learn that `bhead`, `blength`, `delwt`, `fincome`, `gaweeks`, `mheigth`, `momage`, `ppwt`, `wtgain` are related to birthweight(bwt).

*Found the model and show a plot of model residuals against fitted values*

```{r warning=FALSE}
model_1 = lm(bwt ~ bhead + blength + delwt + fincome + gaweeks + mheight + momage + ppwt + wtgain, data = birthweight_df) 

bwt_res_pred_df_1 = 
  birthweight_df %>% 
  modelr::add_predictions(model_1) %>% 
  modelr::add_residuals(model_1) 

  bwt_res_pred_df_1 %>% 
  ggplot(aes(y = resid, x = pred)) + 
  geom_point() +
  geom_smooth(se = F, method = lm)
```

*Cross Validitation to compare the 3 models.*
```{r warning=FALSE}
cv_df = crossv_mc(birthweight_df, 100) %>% 
  mutate(
    model_1 = map(train, ~lm(bwt ~ bhead + blength + delwt + fincome + gaweeks + mheight + momage + ppwt + wtgain, data = .x)),
    model_2 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    model_3 = map(train, ~lm(bwt ~ bhead + blength + babysex + 
               bhead * blength + bhead * babysex + blength * babysex + 
               bhead * blength * babysex, data = .x))
  ) %>% 
  mutate(
    rmse_model_1 = map2_dbl(model_1, test, ~rmse(model = .x, data = .y)),
    rmse_model_2 = map2_dbl(model_2, test, ~rmse(model = .x, data = .y)),
    rmse_model_3 = map2_dbl(model_3, test, ~rmse(model = .x, data = .y)))

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```
Among these 3 models, the model I built above (model_1) has the least rmse in total, which means this model is the better one compared to the other 2 given in the problem.

## Problem 2

```{r warning=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

*Using bootstrap to get the sample*
```{r warning=FALSE}
sample_df = 
  function(df) {
  sample_frac(df, size = 1, replace = TRUE)
}

boot_strap_df = 
  tibble(
    strap_number = 1:5000,
    strap_sample = rerun(5000, sample_df(weather_df))
  )
```

*r^2*
```{r warning=FALSE}
adj_r_square_df = 
  boot_strap_df %>% 
  mutate(
    models = map(.x = strap_sample, ~lm(tmax ~ tmin, data = .x)),
    glance = map(models, broom::glance)
  ) %>% 
  select(strap_number, glance) %>% 
  unnest(glance)

ci_r_square = 
  adj_r_square_df %>% 
  summarize(
    ci_lower = quantile(adj.r.squared, 0.025), 
    ci_upper = quantile(adj.r.squared, 0.975)) %>% 
  round(digits = 2)

adj_r_square_df
adj_r_square_df %>% 
  ggplot(aes(x = adj.r.squared)) +
  geom_density() +
  labs(
        title = "Distribution of R Squared",
        x = "R Squared"
    ) +
  theme(plot.title = element_text(hjust = 0.5))
```
The distribution of `r^2` approximately follows a normal distribution, with a center around 0.91. The 95% confidence interval for r-squared distribution is `r ci_r_square`.

*log(B0 B1) *
```{r warning=FALSE}
log_df = 
    boot_strap_df %>% 
  mutate(
    models = map(.x = strap_sample, ~lm(tmax ~ tmin, data = .x)),
    results = map(models, broom::tidy)
  ) %>% 
  select(strap_number, results) %>% 
  unnest(results) %>% 
  select(term, estimate) %>% 
  pivot_wider(
    names_from = term,
    values_from = estimate
    ) %>% 
  unnest(cols = c(`(Intercept)`, tmin)) %>% 
  rename(B0 = "(Intercept)",
         B1 = tmin) %>% 
  mutate(value = log(B0 * B1))

ci_log = 
  log_df %>% 
  summarize(
    ci_lower = quantile(value, 0.025), 
    ci_upper = quantile(value, 0.975)) %>% 
  round(digits = 2)

log_df
log_df %>% 
  ggplot(aes(x = value)) +
  geom_density() +
  labs(
        title = "Distribution of log(B0 * B1)",
        x = "log(B0*B1)"
    ) +
  theme(plot.title = element_text(hjust = 0.5))
```
The log function plot generally follows a normal distribution, with a center around 2.01. The confidence interval of the log function is `r ci_log`.
